---
title: "Exercise 2: Assignment"
author: "Philipp Kling"
date: "16.06.2020"
output:
  html_document: 
    code_folding: show
    fig_height: 6
    highlight: tango
    theme: spacelab
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---



```{r knitr_init, echo=TRUE, cache=FALSE, warning=FALSE}
library(knitr)

### Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
rm(list = ls())
```

# Preparation

# Assignment 1: Recoginition, loading, extraction


### Recognition

In the Github repository under data/ex2, you will find two files: file_a and file_b. Use a text editor of your choice, inspect each file, and determine what type of data each file includes. 

### File A
Use an appropriate package to load the **file_a** into R. Try the functions associated with the package you used to load the data to get a feeling for the dataset and extract the following information with it. 

 * Extract a list of the IDs of the books.
 * What is the name of the author of the 4th book?

```{r,echo=TRUE}
rm(list = ls())
```


### File B

Use an appropriate package to load the **file_b** into R. 

* What is the email of Mr. Bea?

```{r,echo=TRUE}
rm(list = ls())
```



# Assignment 2: Extract date with Regular Expressions {.tabset .tabset-fade .tabset-pills}

You have the URL to an article of the Guardian. Use regular expressions to extract the publication date of the linked article. Your procedure should be generally applicable to the articles of The Guardian and not only apply to this single link. Use the package __stringr__ and its function __str_extract__ or optionally __str_replace__ to solve this task. (Hint: Slashes and Backslashes must be escaped with two preceding backslashes; e.g. '\ \ /' or '\ \ \'. The package lubridate and its function as_date() are helpful when trying to transform text to dates.)

```{r,echo=TRUE}
url <- "https://www.theguardian.com/society/2018/oct/11/new-law-employers-reveal-race-pay-gap-figures"
# Structure: https://www.theguardian.com/category/year/month/day/headlinetext
result <- lubridate::as_date("2018-10-11")  # The result should look like this.
```


# Assignment 3: Translation

Take the content of file_b used in the first assignment. Transform it into an XML-File format by hand. You can use 
https://www.xmlvalidation.com to validate your attempts.