---
title: "Exercise 4: Assignment"
author: "Philipp Kling"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_folding: show
    fig_height: 6
    highlight: tango
    theme: spacelab
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---



```{r knitr_init, echo=TRUE, cache=FALSE, warning=FALSE}
library(knitr)

### Global options
options(max.print="150")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
rm(list = ls())
```


# Assignment 1: Scrape static webpages

Visit the Wikipedia page for the winners of the Eurovision Song Contest since 1956  (https://en.wikipedia.org/wiki/List_of_Eurovision_Song_Contest_winners). Download the table 'Winners by year' and show which two countries are the countries who won the contest the most frequently.

# Assignment 2: Interact with dynamic pages

**Assignment:** write a script that visits the webpage of the login page of Twitter and tries to log in with the provided credentials.

```{r,echo=T}
url <- "https://www.twitter.com/login"
username <- "testusername"
password <- "testpassword"
```

# Assignment 3: Start an own web scraping attempt

 * Write a short script that evaluates data of a website of your choice and transforms it into a data format R can handle.
 * Describe a way this data could be used in a potential project and provide some summary statistics of variables of interest.
 * Check for alternative ways the website provider offers to access the data (e.g. does the provider have an application programming interface [API; next session]?).
 * Check the robots.txt file of the website. What does it allow, what does it not allow ('Dissallow')?
 * Check for alternative file dumps. Is there any indication that someone already downloaded contents of the website and provided it publicly?
